\documentclass[floatfix,aps,prd,amsmath,amssymb]{revtex4}

\usepackage{epsfig}
\usepackage{color}
\usepackage{graphicx}
\usepackage{float}
\usepackage{listing} %listings
\usepackage{braket}
\usepackage{mathtools}
\providecommand{\e}[1]{\ensuremath{\times 10^{#1}}} %because my scientific notation wouldn't work: Kevin

\begin{document}
\title{HEPP-CPV-project}
\author{John Ronayne (10318997), Kevin Maguire (10318135), Sinead Hales(10318221), Dudley Grant (10275291)}
\date{\today}

\begin{abstract}
\textit{}
\end{abstract}

\maketitle
\pagenumbering{arabic}
 
\input{Introduction}

\input{P_C_CPT}

\input{CP_Violation}

\input{CPV_in_Kaon_system}

%\input{HEPPCPV-CKM}

%\input{CKM_Mechanism}

\input{SCPVsect} %Commented out to avoid clutter.

\begin{appendix}
\section{Appendix}
Difficult calculations in here.
\end{appendix}
 
\begin{thebibliography}{10}
\bibitem{FirstCPV}
``Evidence for the $2 \pi$ Decay of the $K^0_2$ Meson'' - J. H. Christenson, J. W. Cronin, V. L. Fitch, and R. Turlay (1964) Phys. Review letters, vol. 13, issue 4
\bibitem{PDGKaons}
``2013 Review of Particle Physics'' - J. Beringer et al. (Particle Data Group), Phys. Rev. D86, 010001 (2012) and 2013 partial update for the 2014 edition.
\bibitem{Martin+Shaw}
``Particle Physics, Third Edition'' - B.R Martin, G. Shaw, Wiley(2008)  
\bibitem{Christenson_apparatus_ref}
http://large.stanford.edu/courses/2008/ph204/coleman1/  
\bibitem{Measurements_Direct_CPV_Kaons_KTev}
``Measurements of Direct CP Violation, CPT Symmetry, and Other Parameters in the Neutral Kaon System'' - KTeV Collaboration, arXiv:hep-ex/0208007v1 6 Aug 2002
\bibitem{HEPP_intro_Perkins}
``Introduction to High Energy Physics'' - Donald H. Perkins, Cambridge University Press, 4th edition.
\bibitem{DAmbrosio}
``CP violation in Kaon Decays'' - Giancarlo D’Ambrosio and Gino Isidori, arXiv:hep-ph/9611284v1 8 Nov 1996
\bibitem{AsymmetryPic}
``A Measurement of the K(L)-K(s) Mass Difference from the Charge Asymmetry in Semileptonic Kaon Decays'' - Gjesdal, S. et al. Phys.Lett. B52 (1974) 113 Print-74-1358 (CERN)

%SCPV SECTION
\bibitem{SCPV1}
``Spontaneous CP Violation"- Fran\c{c}ois Goffinet, CP3 Seminar, Unit\`{e} de Physique Th\`{e}orique et Math\`{e}matique, U.C.L., December 2003
\bibitem{SCPV2} 
``Spontaneous CP Violation in SUSY" - Yoav Achiman, Physics Letters B, March 2007
\bibitem{SCPV3} 
``A Course in Modern Mathematical Physics" - Peter Szekeres, Cambridge University Press, 2004
\bibitem{SCPV4}
``The Algebra of Grand Unified Theories" - John Baez and John Huerta, Bulletin of the American Mathematical Society, vol 47, May 4 2010
\bibitem{SCPV5}
 ``Group-theoretic Condition for Spontaneous CP Violation" - Howard E. Haber and Ze`ev Surujon, Physical Review D, Volume 86, Issue 7, October 2012
\bibitem{SCPV6}
 ``Field Theory in Particle Physics" Lecture Notes -  Bernard de Wit and Eric Laenen, Universiteit Utrecht, 2009 http://www.staff.science.uu.nl/~wit00103/ftip/Ch11.pdf
\bibitem{SCPV7}
``Kaon Spontaneous CP Violation Reevaluated" - Paul H. Frampton and Masayasu Harada, Physical Review D, Volume 59, Issue 1, March 1998


\end{thebibliography}

\end{document}
